#-----------------------------------------------------------------------
#-----------------------------------------------------------------------
#      Getting and Cleaning Data Project John Hopkins Coursera
#-----------------------------------------------------------------------
#-----------------------------------------------------------------------


#-----------------------------------------------------------------------
# Part 1 - Merges the training and the test sets to create one data set
#-----------------------------------------------------------------------

```{r}
# Read files in ./UCI HAR Dataset
activity_labels <- read.table('./UCI HAR Dataset/activity_labels.txt', 
                              col.names = c('activityId', 'activityName'), quote = "")
features <- read.table('./UCI HAR Dataset/features.txt', 
                       col.names = c('featureId', 'featureName'), quote = "")

# Read files in ./UCI HAR Dataset/test
subject_test <- read.table('./UCI HAR Dataset/test/subject_test.txt', 
                           col.names = c('subjectId'))
X_test <- read.table('./UCI HAR Dataset/test/X_test.txt')
y_test <- read.table('./UCI HAR Dataset/test/y_test.txt')

# Read files in ./UCI HAR Dataset/train
subject_train <- read.table('./UCI HAR Dataset/train/subject_train.txt', 
                            col.names = c('subjectId'))
X_train <- read.table('./UCI HAR Dataset/train/X_train.txt')
y_train <- read.table('./UCI HAR Dataset/train/y_train.txt')

# Combine test data and set column names
colnames(X_test) <- features$featureName
colnames(y_test) <- c('activity')
test_data <- cbind(subject_test, X_test, y_test)

# Combine train data and set column names
colnames(X_train) <- features$featureName
colnames(y_train) <- c('activity')
train_data <- cbind(subject_train, X_train, y_train)

#Combine test and training data
data <- rbind(test_data, train_data)
```

#-----------------------------------------------------------------------
# Part 2 - Extracts only the measurements on the mean and standard 
#          deviation for each measurement.
#-----------------------------------------------------------------------

```{r}
# Extracts features with mean and std [No include meanFreq()]
meanStdData <- data[, c(1, 563, grep(pattern = 'mean\\(\\)|std\\(\\)', x = names(data)))]
# Include the subject ID [col = 1] and activity [col = 563]
```

#-----------------------------------------------------------------------
# Part 3 - Uses descriptive activity names to name the activities in the 
#          data set
#-----------------------------------------------------------------------

```{r}
meanStdData$subjectId <- as.factor(meanStdData$subjectId)

# Change activity column by a factor of 6 levels, with the label the
# same as the activity name
meanStdData$activity <- factor(meanStdData$activity,
                               levels = activity_labels$activityId,
                               labels = activity_labels$activityName)
```

#-----------------------------------------------------------------------
# Part 4 - Appropriately labels the data set with descriptive variable 
#          names.
#-----------------------------------------------------------------------

```{r}
# Remove the () for the mean and std in the measurements names
colnames(meanStdData) <- gsub(pattern = '\\(\\)', replacement = "", x = names(meanStdData))

# Create tidyData file
write.table(meanStdData, file = 'tidyData.txt', row.names = F, quote = F, sep = "\t")
```

#-----------------------------------------------------------------------
# Part 5 - From the data set in step 4, creates a second, independent 
#          tidy data set with the average of each variable for each 
#          activity and each subject.
#-----------------------------------------------------------------------

```{r}
library(dplyr)
meanSdData_sIdAct <- group_by(meanStdData, subjectId, activity) %>% summarise_all(funs(mean))

# Create tidyDataMean file
write.table(meanSdData_sIdAct, file = 'tidyDataMean.txt', 
            row.names = F, quote = F, sep = "\t")
```
